---
title: "coursera-reproducible-research-assignment1"
author: Ziyu Hu
date: September 7, 2020
output: md_document
---
# coursera-reproducible-research-assignment1


## Loading and preprocessing the data

#### 1.load the data
```{r,echo=TRUE}
unzip("./activity.zip")
data <- read.csv("./activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
steps_perday <- aggregate(steps~date,data,sum,na.rm=TRUE)
```
#### 1.make a histogram of the total number of the steps taken each day
```{r}
hist(steps_perday$steps)
```
#### 2.Calculate and report the mean and median total number of steps taken per day
```{r}
mean_steps_perday <- mean(steps_perday$steps)
median_steps_perday <- median(steps_perday$steps)
```
* Mean: `r mean_steps_perday`
* Median: `r median_steps_perday`

## What is the average daily activity pattern?
```{r}
steps_perinterval <- aggregate(steps~interval,data,mean,na.rm=TRUE)
```
#### 1.make a time series plot
```{r}
plot(steps~interval,steps_perinterval,type="l")
```
#### 2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_step_interbal <- steps_perinterval[which.max(steps_perinterval$steps),]$interval
```
* Most steps at: `r max_step_interbal`

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset
```{r}
totalValuesMissings <- sum(is.na(data$steps))
```
* Number of missing values: `r totalValuesMissings`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
we’ll fill in all the missing values in the dataset with the mean per interval. Here’s the function that will return, for a particular interval, the mean value
```{r}
getMeanStepsPerInterval<-function(interval){
    steps_perinterval[steps_perinterval$interval==interval,]$steps
}
```

##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data_noNA<-data
for(i in 1:nrow(data)){
    if(is.na(data[i,]$steps)){
        data_noNA[i,]$steps <- getMeanStepsPerInterval(data_noNA[i,]$interval)
    }
}
```
##### 4. Make a histogram of the total number of steps taken each day 
```{r}
data_noNA_sum <- aggregate(steps ~ date, data=data_noNA, sum)
hist(data_noNA$steps)
```   
##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
mean <- mean(data_noNA_sum$steps)
median <- median(data_noNA_sum$steps)
```
* Mean: `r mean`
* median: `r median`

## Are there differences in activity patterns between weekdays and weekends?

##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
library(ggplot2)
library(dplyr)
library(lubridate)
```
```{r}
day_of_week <- data_noNA %>%
  mutate(
    date = ymd(date),
    weekday_or_weekend = case_when(wday(date) %in% 2:6 ~ "Weekday",
                                   wday(date) %in% c(1,7) ~ "Weekend")
  ) %>% select(-date) %>%
  group_by(interval, weekday_or_weekend) %>%
  summarise(
    steps = mean(steps)
  )
```

##### 2. Make a panel plot containing a time series plot
```{r}
ggplot(day_of_week, aes(interval, steps)) + 
  geom_line() + 
  facet_wrap(~weekday_or_weekend, nrow = 2) +
  xlab("5-Minute intervals") + 
  ylab("Average number of steps")

```

